Using a Highly Dynamic Language for Development

### 2021-10-10 gic2021-highly-dynamic/paper.tex
Minor revisions

\begin{abstract}
  Games face an interesting challenge. They require rapid development, are highly interactive, and pose hard real-time performance constraints. While smaller games these days have also been developed in dynamic languages such as Python or Lua, traditionally engines are still written in static languages like C++ and C, with an additional scripting language on top to handle gameplay mechanics. Common Lisp offers an environment that's both dynamic and performant enough to allow for a full stack game development system that is highly favourable to fast iteration and modular design.
\end{abstract}
\begin{keywords}
  Common Lisp, game development, dynamic languages, object orientation
\end{keywords}

@@

\section{Introduction}
Video games pose an interesting engineering challenge. They are highly dynamic in their nature, as users can perform various, sometimes far-reaching changes to the program at any time, and yet they must remain responsive under hard real-time constraints. Additionally, the development of games itself is highly dynamic, as changes to the game require constant testing and refinement. Long pauses between making a change and being able to properly evaluate its effects can gravely discourage testing, which leads to a much worse product.

@@

\begin{listing}[h]
\begin{minted}[fontsize=\scriptsize]{common-lisp}
(defmethod handle :before ((ev tick) (object moving-entity)))
\end{minted}
\caption{The updated second method definition}
\label{lst:updated method}
\end{listing}

@@

In our UI toolkit Alloy we use this trick extensively to avoid the allocation of colour values, sizes, and dimensional units.

@@

While these features offer great amounts of flexibility, there is a performance cost to be paid. Many of these performance penalties can be avoided or circumvented with careful planning and design, but often by sacrificing some of the dynamism. These optimisations also come at a greater development cost.

>>>>

\begin{abstract}
  Games face an interesting challenge. They require rapid development, are highly interactive, and pose soft real-time performance constraints. While smaller games these days have also been developed in dynamic languages such as Python or Lua, traditionally engines are still written in static languages like C++ and C, with an additional scripting language on top to handle gameplay mechanics. Common Lisp offers an environment that's both dynamic and performant enough to allow for a full stack game development system that is highly favourable to fast iteration and modular design.
\end{abstract}
\begin{keywords}
  Common Lisp, game development, dynamic languages, object orientation
\end{keywords}

@@

\section{Introduction}
Video games pose an interesting engineering challenge. They are highly dynamic in their nature, as users can perform various, sometimes far-reaching changes to the program at any time, and yet they must remain responsive under soft real-time constraints. Additionally, the development of games itself is highly dynamic, as changes to the game require constant testing and refinement. Long pauses between making a change and being able to properly evaluate its effects can gravely discourage testing, which leads to a much worse product.

@@

\begin{listing}[h]
\begin{minted}[fontsize=\scriptsize]{common-lisp}
(defmethod handle :before ((ev tick) (object physics-object)))
\end{minted}
\caption{The updated second method definition}
\label{lst:updated method}
\end{listing}

@@

In Alloy, our UI toolkit, we use this trick extensively to avoid the allocation of colour values, sizes, and dimensional units.

@@

While these features offer great amounts of flexibility, there is a performance cost. Many of these performance penalties can be avoided or circumvented with careful planning and design, but often by sacrificing some of the dynamism. These optimisations also come at a greater development cost.
